---
title: "Reproducible Research: Peer Assessment 1 By Mahmood Karimi"
output: 
  html_document:
    keep_md: true
---
This project is about ananlysis of step-tracking data for two month of a person. We show you how has been the activity of the person during these two months.

## Loading and preprocessing the data
```{r 1st chunck}
unzip("./activity.zip", exdir = "./data")
df <- read.csv("./data/activity.csv")
```


## What is mean total number of steps taken per day?
```{r 2nd chunk}
agd <- aggregate(df$steps, by=list(df$date), FUN=sum)
names(agd) <- c("date", "sumsteps")
par(col="red", las=3)
plot(agd$date, agd$sumsteps, type = "l", main = "#steps per day", xlab = "", 
     ylab = "#steps")
abline(h=mean(agd$sumsteps, na.rm=TRUE))
summary(agd$sumsteps)
boxplot(agd$sumsteps)
```


## What is the average daily activity pattern?
```{r 3rd chunk}
tsa <- ts(df)
plot.ts(tsa, main="Activity pattern")
agd <- aggregate(df$steps, by=list(df$interval), FUN=mean, na.rm=TRUE)
names(agd) <- c("interval", "meansteps")
t <- agd[agd$meansteps==max(agd$meansteps),]
```

The `r t$interval` interval contrains maximum number of steps on average.


## Imputing missing values
The strategy for imputing has been different. For the day that didn't have any similar equivalent in another month the mean of total days is used.  
For the days that had similar equivalent days, the corresponding equivalent day is used.

Days with NA values:
```{r 4th chunk}
# Finding the days with NA values
na_steps <- table(df[is.na(df$steps), "date"])
na_df <- as.data.frame(na_steps)
na_df[na_df$Freq > 0,]

# Define two functons to impute NA values
copy_date <- function(df1, from_date, to_date) {
        # df1[to_date, steps] <- df1[from_date, steps]
        t1df <- df1[df1$date != to_date,]
        t2df <- df1[df1$date == from_date,]
        t2df$date <- to_date
        df1 <- rbind(t1df, t2df)
        df1
}

set_avg <- function(df1, dt) {
        # copy the avg of steps to steps values in dt
        df2 <- aggregate(df1$steps, by=list(df1$interval), FUN=mean, na.rm=TRUE)
        names(df2) <- c("interval", "steps")
        df2$steps <- round(df2$steps)
        t1df <- df1[df1$date != dt,]
        t2df <- df1[df1$date == dt,]
        t2df[,c("interval", "steps")] <- df2
        df1 <- rbind(t1df, t2df)
        df1
}

# Imputing the NA values
df <- set_avg(df, "2012-10-01")
df <- copy_date(df, "2012-11-08","2012-10-08")
df <- copy_date(df, "2012-10-01","2012-11-01")
df <- copy_date(df, "2012-10-04","2012-11-04")
df <- copy_date(df, "2012-10-09","2012-11-09")
df <- copy_date(df, "2012-10-10","2012-11-10")
df <- copy_date(df, "2012-10-14","2012-11-14")
df <- copy_date(df, "2012-10-30","2012-11-30")

# Drawing the result
agd <- aggregate(df$steps, by=list(df$date), FUN=sum)
names(agd) <- c("date", "sumsteps")
par(col="red", las=3)
plot(agd$date, agd$sumsteps, type = "l", main = "#steps per day", xlab = "", 
     ylab = "#steps")
abline(h=mean(agd$sumsteps))
summary(agd$sumsteps)
```


## Are there differences in activity patterns between weekdays and weekends?
```{r 5th chunk}
df$dow <- weekdays(as.Date(df$date))
df$holiday <- (df$dow == "ÔäÈå" | df$dow == "íßÔäÈå")
df_we <- df[df$holiday,]
df_wd <- df[!df$holiday,]
agd_we <- aggregate(df_we$steps, by=list(df_we$interval), FUN=mean)
agd_wd <- aggregate(df_wd$steps, by=list(df_wd$interval), FUN=mean)
names(agd_we) <- c("interval", "meansteps_we")
names(agd_wd) <- c("interval", "meansteps_wd")
agd <- merge(agd_wd, agd_we, by="interval")
par(mfrow=c(1,2))
plot(agd$interval, agd$meansteps_wd, main="Average #steps", xlab = "Time", ylab = "Weekdays")
plot(agd$interval, agd$meansteps_we, main="Average #steps", xlab = "Time", ylab = "Weekends")
```